The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-24 writing/draft3/draft17.tex
writing

Besides our main model, we estimated the same specification for two additional subsamples of the Kenyan counties. In particular, we estimated the model for the subsample of the ASAL counties and for the subsample of the non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with the significantly different climate. The estimates of all three models are summarised in table~\ref{MainEst}. The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of the ASAL counties and the last two columns describe the estimates based on the subsample of the non-ASAL counties. All predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance. As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance (ANOVA) in table~\ref{MainEst}. The F-tests were conducted using the marginal rather than the sequential sum of squares. That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable. In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors. The main advantage of the type~\textit{III} ANOVA is that the test statistics do not depend on the order in which the effects are entered in the model.

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects on yield are multiplicative. The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix. The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised. The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.

In the following, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.

@@

The estimates of the model based on all the Kenyan counties are summarised in the first two columns of table~\ref{MainEst}. Consistently with our assumptions, seasonal precipitation and average temperature have significant effect on yields. As we expected, the effect of temperature is negative. The linear term of precipitation is positive while the squared term is negative and as we can see in figure~\ref{MarEff1}(a), the marginal effect of precipitation is hill-shaped

>>>>

Besides our main model, we estimated the same specification for two additional subsamples of the Kenyan counties. In particular, we estimated the model for the subsample of the ASAL counties and for the subsample of the non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with the significantly different climate. The estimates of all three models are summarised in table~\ref{MainEst}. The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of the ASAL counties and the last two columns describe the estimates based on the subsample of the non-ASAL counties. All predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance. As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance (ANOVA) in table~\ref{MainEst}. The F-tests were conducted using the marginal rather than the sequential sum of squares. That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable. In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors. The main advantage of the type~\textit{III} ANOVA is that the test statistic does not depend on the order in which the effects are entered in the model.

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects are multiplicative. The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix. The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised. If the values of the marginal effects are smaller than one, the effect of yield is negative while marginal effects higher than one can be interpreted as positive effect. The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.

In the following few paragraphs, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.

@@

The estimates of the model based on all the Kenyan counties are summarised in the first two columns of table~\ref{MainEst}. Consistently with our assumptions, seasonal precipitation and average temperature have significant effect on yields. As we expected, the effect of temperature is negative. The linear term of precipitation is positive while the squared term is negative and as we can see in figure~\ref{MarEff1}(a), the marginal effect of precipitation is hill-shaped. That is, the effect of precipitation is negative and increasing if its amount is below mean then the effect further increases with the amount of precipitation until it reaches $1.403$ times its standard deviation. At this point the effect starts to decrease with increasing precipitation and it became negative again once the amount reaches $2.805$ times its standard deviation.
